---
title: "Arkansas Weekly Weather & Soil Water Report"
subtitle: "Week of June 23, 2025"
author: "University of Arkansas Division of Agriculture"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    page-layout: full
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(data.table)
library(ggplot2)

# Load configuration and utilities
source("../code/00-config.R")
source("../code/utils/visualization-theme.R")

# Load processed data
conditions <- readRDS(file.path(PATH_PROCESSED_DATA, "current-conditions.rds"))
soil_water <- readRDS(file.path(PATH_PROCESSED_DATA, "soil-water-status.rds"))

# Try to load optional data
historical_stats <- tryCatch(
  readRDS(file.path(PATH_PROCESSED_DATA, "historical-stats.rds")),
  error = function(e) NULL
)
forecasts <- tryCatch(
  readRDS(file.path(PATH_PROCESSED_DATA, "forecast-trends.rds")),
  error = function(e) NULL
)
```

## Executive Summary

**Report Date:** `r format(Sys.Date(), "%B %d, %Y")`
**Week Covered:** June 16–22, 2025
**Report Timezone:** Central Daylight Time (CDT)

### Key Highlights

- **Temperature:** [Summary of temperature for the week]
- **Precipitation:** [Summary of rainfall and comparison to normal]
- **Soil Water:** [Overall state across research stations]
- **Alerts:** [Any frost risks, drought stress, or other alerts]

---

## Statewide Overview

### Temperature Analysis

```{r temp-summary}
temp_summary <- conditions %>%
  filter(date >= REPORT_DATE - 7, date <= REPORT_DATE) %>%
  summarise(
    tmax_mean = mean(tmax, na.rm = TRUE),
    tmin_mean = mean(tmin, na.rm = TRUE),
    tmean_mean = mean(tmean, na.rm = TRUE),
    tmax_max = max(tmax, na.rm = TRUE),
    tmin_min = min(tmin, na.rm = TRUE),
    .by = station
  )

knitr::kable(temp_summary,
  caption = "Temperature Summary (Last 7 Days)",
  digits = 1
)
```

### Precipitation

```{r precip-plot}
precip_weekly <- conditions %>%
  filter(date >= REPORT_DATE - 7, date <= REPORT_DATE) %>%
  group_by(station) %>%
  summarise(
    total_precip = sum(precip, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(total_precip))

ggplot(precip_weekly, aes(x = reorder(station, total_precip), y = total_precip)) +
  geom_col(fill = "#2ca25f", alpha = 0.7) +
  geom_text(aes(label = sprintf("%.1f mm", total_precip)),
            vjust = -0.5, size = 3) +
  labs(
    title = "Total Precipitation by Station (Last 7 Days)",
    x = "Research Station",
    y = "Precipitation (mm)"
  ) +
  theme_ar_weather() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

## Historical Benchmarking

### How This Week Compares

```{r benchmark}
if (!is.null(historical_stats)) {
  bench_table <- historical_stats %>%
    select(station, week, tmax_pctl, tmin_pctl, precip_pctl,
           tmax_anomaly, tmin_anomaly, precip_anomaly)

  knitr::kable(bench_table,
    caption = "Historical Percentiles & Anomalies",
    digits = 1
  )
} else {
  cat("*Historical benchmark data not yet available*")
}
```

**Interpretation:**
- **75th percentile and above:** Unusually warm/wet (rare for this time of year)
- **25th–75th percentile:** Near-normal conditions
- **Below 25th percentile:** Unusually cool/dry (rare for this time of year)

---

## Soil Water Status

### Current Relative Soil Water

```{r soil-water-map}
soil_latest <- soil_water %>%
  filter(date == max(date)) %>%
  arrange(desc(rel_soil_water))

ggplot(soil_latest, aes(x = reorder(station, rel_soil_water),
                        y = rel_soil_water,
                        fill = rel_soil_water)) +
  geom_col() +
  scale_fill_gradient(low = "#8B4513", mid = "#DAA520", high = "#90EE90",
                      limits = c(0, 100),
                      name = "Rel. Soil Water (%)") +
  geom_hline(yintercept = STRESS_THRESHOLD * 100,
             linetype = "dashed", color = "red", size = 1) +
  geom_text(aes(label = sprintf("%.0f%%", rel_soil_water)),
            vjust = -0.5, size = 3) +
  labs(
    title = "Relative Soil Water Status by Station",
    subtitle = "Red line indicates stress threshold",
    x = "Research Station",
    y = "Relative Soil Water (%)"
  ) +
  theme_ar_weather() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")
```

### Soil Water Trend (Last 30 Days)

```{r soil-water-trend}
soil_trend <- soil_water %>%
  filter(date >= REPORT_DATE - 30) %>%
  arrange(station, date)

ggplot(soil_trend, aes(x = date, y = rel_soil_water, color = station)) +
  geom_line(size = 0.8) +
  geom_hline(yintercept = STRESS_THRESHOLD * 100,
             linetype = "dashed", color = "red", alpha = 0.5) +
  facet_wrap(~station, scales = "free_y") +
  labs(
    title = "Soil Water Trend (Last 30 Days)",
    x = "Date",
    y = "Relative Soil Water (%)"
  ) +
  theme_ar_weather() +
  theme(legend.position = "none")
```

---

## Growing Degree Days (GDD)

```{r gdd}
gdd_summary <- conditions %>%
  filter(date >= as.Date("2025-04-01")) %>%
  group_by(station) %>%
  summarise(
    cum_gdd = max(cum_gdd, na.rm = TRUE),
    days_since_april = as.numeric(max(date) - as.Date("2025-04-01")),
    .groups = "drop"
  ) %>%
  mutate(
    gdd_per_day = cum_gdd / days_since_april
  )

knitr::kable(gdd_summary %>% arrange(desc(cum_gdd)),
  caption = "Cumulative Growing Degree Days (Since April 1, 2025)",
  digits = 1
)
```

**GDD Milestones:**
- 100 GDD = Emergence
- 400–600 GDD = V4–V6 growth
- 900–1200 GDD = Flowering begins
- 1500–2000 GDD = Physiological maturity

---

## 7–10 Day Forecast

```{r forecast}
if (!is.null(forecasts)) {
  forecast_summary <- forecasts %>%
    group_by(station) %>%
    summarise(
      temp_trend = mean(temp_trend, na.rm = TRUE),
      frost_risk_days = sum(frost_risk, na.rm = TRUE),
      precip_prob_avg = mean(precip_prob, na.rm = TRUE),
      stress_days = sum(stress_signal, na.rm = TRUE),
      .groups = "drop"
    )

  knitr::kable(forecast_summary,
    caption = "7-Day Forecast Summary",
    digits = 1
  )
} else {
  cat("*Forecast data not yet available*")
}
```

### Frost Risk Alert
```{r frost-check}
if (!is.null(forecasts)) {
  frost_risk <- forecasts %>%
    filter(frost_risk == 1) %>%
    distinct(station, forecast_date) %>%
    arrange(forecast_date)

  if (nrow(frost_risk) > 0) {
    cat("**⚠️ FROST RISK DETECTED:**\n\n")
    for (i in seq_len(nrow(frost_risk))) {
      cat(sprintf("- %s: %s\n",
                  frost_risk$station[i],
                  format(frost_risk$forecast_date[i], "%B %d")))
    }
  } else {
    cat("**✓ No frost risk in 7-day forecast**")
  }
} else {
  cat("*Forecast data not available*")
}
```

---

## Station Details

### Fayetteville

**Location:** 36.07°N, 94.17°W | Elevation: 395 m

```{r fay-detail}
fay_data <- conditions %>%
  filter(station == "Fayetteville", date >= REPORT_DATE - 7) %>%
  arrange(date) %>%
  select(date, tmax, tmin, precip, cum_gdd)

knitr::kable(fay_data,
  caption = "Fayetteville: Last 7 Days",
  digits = 1
)
```

*Add additional station sections as needed (Marianna, Pine Bluff, Rohwer, Hope)*

---

## Crop Management Alerts

### Water Stress Risk
```{r stress-alert}
stress_stations <- soil_water %>%
  filter(date == max(date), rel_soil_water < (STRESS_THRESHOLD * 100)) %>%
  arrange(rel_soil_water)

if (nrow(stress_stations) > 0) {
  cat("**⚠️ WATER STRESS RISK:**\n\n")
  for (i in seq_len(nrow(stress_stations))) {
    cat(sprintf("- %s: %.0f%% available water\n",
                stress_stations$station[i],
                stress_stations$rel_soil_water[i]))
  }
  cat("\n*Consider irrigation if available*\n")
} else {
  cat("**✓ No water stress alert this week**")
}
```

### Temperature Extremes
```{r temp-alert}
extreme_temp <- conditions %>%
  filter(date >= REPORT_DATE - 7) %>%
  filter(tmax > 35 | tmin < 0)

if (nrow(extreme_temp) > 0) {
  cat("**⚠️ TEMPERATURE EXTREMES DETECTED:**\n\n")
  print(unique(extreme_temp %>% select(date, station, tmax, tmin)))
} else {
  cat("**✓ No temperature extremes this week**")
}
```

---

## Data Quality & Methods

### Data Sources
- **Weather:** Iowa Environmental Mesonet (IEM), Open-Meteo API
- **Forecasts:** Open-Meteo 7–10 day model
- **Soil Properties:** USDA SSURGO soil survey database

### Quality Assurance
- Automated outlier detection and removal
- Gap-filling using linear interpolation (max 3-day gaps)
- Cross-validation with nearby stations
- Visual inspection of plots

### Model Methods
- **Soil Water:** Simple bucket model (available water capacity)
- **ET Rate:** Hargreaves equation (temperature, solar radiation)
- **GDD:** Base temperature 10°C (50°F), no upper limit
- **Historical:** 20-year baseline (1995–2020)

---

## Download Data

All data from this report is available for download as CSV files:

- [Weather Data](../data-downloads.qmd#current-weeks-data)
- [Soil Water Data](../data-downloads.qmd#current-weeks-data)
- [Forecast Data](../data-downloads.qmd#current-weeks-data)

See [Data Downloads](../data-downloads.qmd) page for historical archives and complete data dictionary.

---

## Contact & Questions

**Lab:** University of Arkansas Division of Agriculture
**Project Lead:** Dr. Eduardo Ellis
**Email:** [eelli@uark.edu](mailto:eelli@uark.edu)
**GitHub:** [ar-weekly-weather-soil-reports](https://github.com/your-lab/ar-weekly-weather-soil-reports)

---

**Report Generated:** `r format(Sys.time(), "%B %d, %Y at %H:%M %Z")`

See [Home](../index.qmd) | [Reports Archive](archive.qmd) | [Lab Updates](../lab-updates.qmd)
