---
title: "Arkansas Weekly Soil Water Status Report"
subtitle: "Week of June 16, 2025 to June 22, 2025"
date: "June 26, 2026"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    fig-width: 10
    fig-height: 6
params:
  data_file: null
  week: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
library(ggplot2)
library(scales)
library(readr)

# Load data
data <- read_csv(params$data_file, show_col_types = FALSE)

# Extract week info
year <- unique(data$year)
week <- unique(data$week)
week_start <- unique(data$week_start)[1]
week_end <- unique(data$week_end)[1]
n_cells <- nrow(data)
```

## Executive Summary

This report summarizes soil water holding capacity (SWHC) conditions across 59 Arkansas cropland grid cells for week **W24** (June 16, 2025 – June 22, 2025).

**Key Metrics:**
- **6-inch depth:** Mean = 24.5 mm (range: 19.1–30.4 mm)
- **12-inch depth:** Mean = 53.8 mm (range: 47.5–59.3 mm)
- **24-inch depth:** Mean = 108.6 mm (range: 102.1–115.7 mm)

### Interpretation
Soil water availability increases with depth (expected for Arkansas soils). All depths show realistic variation across the spatial domain.

---

## Spatial Distribution

### 6-Inch Depth Soil Water

```{r map-6in}
ggplot(data, aes(x = x, y = y, color = swhc_6in_mean)) +
  geom_point(size = 3, alpha = 0.7) +
  scale_color_viridis_c(option = "plasma", name = "SWHC (mm)") +
  labs(
    title = "Soil Water Holding Capacity - 6-Inch Depth",
    subtitle = paste("Week", week, "–", year),
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "right"
  )
```

### 12-Inch Depth Soil Water

```{r map-12in}
ggplot(data, aes(x = x, y = y, color = swhc_12in_mean)) +
  geom_point(size = 3, alpha = 0.7) +
  scale_color_viridis_c(option = "viridis", name = "SWHC (mm)") +
  labs(
    title = "Soil Water Holding Capacity - 12-Inch Depth",
    subtitle = paste("Week", week, "–", year),
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "right"
  )
```

### 24-Inch Depth Soil Water

```{r map-24in}
ggplot(data, aes(x = x, y = y, color = swhc_24in_mean)) +
  geom_point(size = 3, alpha = 0.7) +
  scale_color_viridis_c(option = "mako", name = "SWHC (mm)") +
  labs(
    title = "Soil Water Holding Capacity - 24-Inch Depth",
    subtitle = paste("Week", week, "–", year),
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "right"
  )
```

---

## Distribution Statistics

### Depth-Wise Comparison

```{r depth-comparison}
# Prepare data for comparison
depth_stats <- data.frame(
  Depth = c("6-inch", "12-inch", "24-inch"),
  Mean = c(
    mean(data$swhc_6in_mean, na.rm = TRUE),
    mean(data$swhc_12in_mean, na.rm = TRUE),
    mean(data$swhc_24in_mean, na.rm = TRUE)
  ),
  Min = c(
    min(data$swhc_6in_min, na.rm = TRUE),
    min(data$swhc_12in_min, na.rm = TRUE),
    min(data$swhc_24in_min, na.rm = TRUE)
  ),
  Max = c(
    max(data$swhc_6in_max, na.rm = TRUE),
    max(data$swhc_12in_max, na.rm = TRUE),
    max(data$swhc_24in_max, na.rm = TRUE)
  )
)

ggplot(depth_stats, aes(x = Depth, y = Mean, fill = Depth)) +
  geom_col(alpha = 0.7) +
  geom_errorbar(aes(ymin = Min, ymax = Max), width = 0.2, color = "black") +
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(
    title = "Soil Water Holding Capacity by Depth",
    subtitle = "Mean with min/max range",
    y = "SWHC (mm)",
    x = "Soil Depth"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 11)
  )
```

### Distribution by Depth

```{r distribution-boxplot}
# Prepare long format for boxplot
data_long <- data %>%
  select(cellid, swhc_6in_mean, swhc_12in_mean, swhc_24in_mean) %>%
  tidyr::pivot_longer(
    cols = starts_with("swhc_"),
    names_to = "Depth",
    values_to = "SWHC_mm"
  ) %>%
  mutate(Depth = factor(
    gsub("swhc_|_mean", "", Depth),
    levels = c("6in", "12in", "24in")
  ))

ggplot(data_long, aes(x = Depth, y = SWHC_mm, fill = Depth)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 2) +
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(
    title = "Distribution of Soil Water Across Grid Cells",
    subtitle = paste("Week", week, "–", year),
    y = "SWHC (mm)",
    x = "Soil Depth"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold")
  )
```

---

## Data Summary Table

```{r summary-table}
summary_table <- data %>%
  select(cellid, x, y, swhc_6in_mean, swhc_12in_mean, swhc_24in_mean) %>%
  head(20) %>%
  mutate(across(starts_with("swhc"), round, 1))

knitr::kable(
  summary_table,
  col.names = c("Cell ID", "Longitude", "Latitude", "6in (mm)", "12in (mm)", "24in (mm)"),
  caption = paste("Sample of first 20 cells (out of", n_cells, "total)")
)
```

---

## Methods

**Data Source:** APSIM Next Generation crop simulations (baseline MG4 soybean scenario)

**Processing:**
1. Grid simulations run for each cell with local weather and soil profiles
2. Soil water metrics extracted daily (swhc_6in, swhc_12in, swhc_24in)
3. Weekly aggregation: mean, min, max per depth per cell
4. Report generated automatically via Quarto

**Time Window:** `r format(week_start, "%Y-%m-%d")` to `r format(week_end, "%Y-%m-%d")`

**Grid Cells:** `r n_cells` cultivated cells across Arkansas

---

## Technical Details

- **Report Generated:** `r Sys.time()`
- **R Version:** `r R.version$version.string`
- **Packages:** dplyr, ggplot2, readr, tidyr

---

**Questions?** Contact the AR Crop Modeling program
**Data:** Available in `data/outputs/soil-water-status-YYYY-WXX.csv`
